---
title: "Practical Machine Learning"
output: html_document
---
## COURSE PROJECT:WRITE-UP

Firstly, we read the training and the test sets into R.

```{r}
training <- read.csv("pml-training.csv", na.strings = c("", "NA"))
testing <- read.csv("pml-testing.csv")
```

Secondly, we check which columns have NAs as well as the percentage of the NAs in those columns.
```{r}
AA <- rep(NA, ncol(training))

for (i in 1:ncol(training)){
     AA[i] <- sum(is.na(training[,i]))/nrow(training)
}

table(AA)
```
In this model we decide to completely disregard the columns in which the percentage of the NAs is more than 80%. This decision leaves us with columns which contain no NAs in this particular example and it is executed with the command below. The new training set is called training2.   

```{r}
training2 <- training[,which(AA == 0)]
```

After that we check whether any of the variables has near zero variance.
```{r}
library(caret)
nsv <- nearZeroVar(training2, saveMetrics = TRUE)
nsv
```


We see that the 6th column has near zero variance. We would like to exlude it as keeping it as a predictor does not add any value. Moreover, the first column represents the row number and it should be excluded as well as done in the following.

```{r}
training2 <- training2[,-c(1,6)]
```

For the method we choose "gbm" with the 10-fold cross validation.
```{r}
set.seed(33833)
model.fit <- train(classe ~ ., method = "gbm", trControl = trainControl(method = "cv", number = 10), data = training2, verbose = FALSE)
```
Below you can see the value of Kappa as in sample error. The out of sample error will bigger than that.
```{r}
confusionMatrix(predict(model.fit, training2), training2$classe)
```

Having built our model we predict the test cases and create the text files.
```{r}
answers <- as.character(predict(model.fit, testing))


pml_write_files = function(x){
       n = length(x)
       for(i in 1:n){
             filename = paste0("problem_id_",i,".txt")
             write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
         }
}
pml_write_files(answers)

```





